Cyclone: High Availability for Persistent Key Value Stores

### 2017-12-23 paper/paper.tex
Updates.

Although the work described in the previous section significantly boosts the
network performance of replication in Cyclone, it still does not come close to
saturating the capabilities of even the commodity NICs that we use. This is
because NICs today encapsulate significant parallelism in terms of
multiple send and receive queue pairs. Exploiting this capability to improve
throughput requires us to remove the bottleneck of a single sequential log in
RAFT. To this end, we extend the implementation described thus far to run
multiple copies of the RAFT consensus protocol each maintaining and replicating
its own two level log. We refer to these logs as physical logs. All these
instances however exist in the same process address space as the key value store
application itself and therefore manipulate the same application. The number of 
instances is a fixed property of the Cyclone service and cannot be changed after 
startup.

@@

Multiple physical logs in Cyclone operate in a shared nothing manner by
partitioning the NVM and SSD space evenly between them and by allocating
dedicated NIC queue pairs to each instance of RAFT. This works well because the
memory hierarchy including the NVM and the SSD efficiently support concurrent
operations. The only synchronization necessary is when doing reads or writes to
the single shared key value store.  The level of concurrency therefore is
constrained only by the concurrency available in the software architecture of
the persistent key value store itself, for which good designs
exist~\cite{flodb}.

\subsection{Ganged Operations}
Using multiple instances of RAFT leads to a serious problem with batched operations, 
since we need to split up the batch into a mini-batch for each of the physical logs. There
is however no guarantee that the request for a mini-batch will succeed breaking the atomicity
requirement for batched updates. The obvious and simple solution is to do two phase commit across
the logs, ensuring all or nothing semantics. However this was unacceptable to some customers who
pointed out that ensuring that an external co-ordinator implementing two phase commit does not itself
fail requires running a separate replicated service - with its attendant resource and reliability headaches.
We were therefore challenged to come up with a solution that did not require two phase commit.

Our final solution is based on two observations. First, we \emph {do not} require that either all mini-batches
icommit to their raft logs or none do, What we actually requires is that either all are \emph{applied} to Rocksdb or 
none of them are. Put another way we should not apply a mini-batch from a ganged operation if any of the other mini-batches
have failed to commit to their RAFT log. This means that the client can be stateless. The second observation is that
checking whether the other minibatches have committed is easy because the different RAFT instances can communicate 
through shared memory on the same machines unlike the more general case of distributed participants in two phase 
commit.

We now describe our solution called ``ganged operations" in Cyclone. 
The first key problem we need to deal with is that the leaders for the different
RAFT instances to which the keys map can be located on different machines. To
avoid having a client co-ordinate different machines for a single operation, we
constrain the RAFT leader election algorithm for the different RAFT instances to
converge to the same machine. We do this by triggering re-elections for a RAFT
instance other than that of the first physical log until it is on the same
machine as the \emph{first} physical log. This process is therefore resilient to
failed machines. RAFT requires that the leader have the most up to date log. To
ensure that the process converges we therefore ensure that the currently elected
leader brings a majority quorum up to date before triggering a reelection. We
also do not accept any requests from clients until all leaders are
co-located. We assume that no network partitions occur \emph{within a machine}
i.e. some RAFT instances are able to send and receive packets, while others are
not, a possibility we discount due to our single process design.

Next, we need to simultaneously inject the ganged operation into all
participating physical logs on the machine hosting the leaders. Clients always
dispatch batched writes to a fixed co-ordinating physical log/RAFT instance
(henceforth called the co-ordinator), which is then responsible for forwarding
the request to the participating logs. We make use of packet cloning primitives
to avoid making physical copies of the packet, generating indirect references
instead. The co-ordinator also adds a ``nonce'' - a unique timestamp to the
packet. In addition, the co-ordinator adds a unique view number to the packet
containing the term numbers of all the participating RAFT instances (read from
shared memory of the co-located leaders).  The event that applies a ganged
operation is described in the pseudocode of Figure~\ref{fig:ganged_ops}. We
defer the discussion of how we generate the nonce to later in this section. We
also assume that a unique barrier is allocated in shared memory for each ganged
operation. We discuss later how this is done without using dynamic allocation.

The complexity in Figure~\ref{fig:ganged_ops} arises from the need to handle
both shared memory concurrency and failure during replication. Without failure,
Figure~\ref{fig:ganged_ops} is straightforward. A barrier is executed on all
RAFT instances. Once replication is complete on all participating physical logs,
a state indicated by all necessary bits being set in the barrier mask, the RAFT
instances simultaneously execute the ganged operation.

@@

The assumption in Figure~\ref{fig:ganged_ops} is that each ganged operation is
mapped to a unique barrier. We achieve this by using a fixed piece of memory
owned by the co-ordinator to hold the barrier and write the nonce to it in order
to indicate that the barrier is active for the corresponding ganged
operation. Participants watch for the nonce to know when to execute the
ganged operation barrier in Figure~\ref{fig:ganged_ops}, while also monitoring
the leader's published view to detect the case where the ganged operation fails
to replicate on the co-ordinator's physical log.

>>>>

Although the work described in the previous section significantly boosts the
network performance of replication in Cyclone, it still does not come close to
saturating the capabilities of even the commodity NICs that we use. This is
because NICs today encapsulate significant parallelism in terms of
multiple send and receive queue pairs. Exploiting this capability to improve
throughput requires us to remove the bottleneck of a single sequential log in
RAFT. To this end, we extend the implementation described thus far to run
multiple copies of the RAFT consensus protocol each maintaining and replicating
its own two level log. We refer to these logs as physical logs. All these
instances however exist in the same process address space as the key value store
application itself and therefore manipulate the same application.

@@

Multiple physical logs in Cyclone operate in a shared nothing manner
by partitioning the NVM and SSD space evenly between them and by
allocating dedicated NIC queue pairs to each instance of RAFT.  The
only synchronization necessary is when doing reads or writes to the
single shared key value store - something already provided by most
key-value stores including Rocksdb.  The level of concurrency
therefore is constrained only by the concurrency available in the
software architecture of the persistent key value store itself, for
which good designs exist~\cite{flodb}.

\subsection{Ganged Operations}
Using multiple instances of RAFT leads to a serious problem with
batched operations, since we need to split up the batch into a
mini-batch for each of the physical logs. There is no guarantee that
the request for a mini-batch will succeed breaking the atomicity
requirement for batched updates. The obvious and simple solution is to
do two phase commit across the logs, ensuring all or nothing
semantics. However this was unacceptable to some use-cases of Cyclone
as it requires ensuring that an external co-ordinator implementing two
phase commit does not itself fail. This requires running a separate
replicated service - with its attendant resource and reliability
headaches - too high a price to pay for the performance benefits of
Cyclone.  We therefore had to come up with a solution that did not
require two phase commit.

Our solution is based on two observations. First, we \emph {do not}
require that either all mini-batches icommit to their raft logs or
none do. What we actually requires is that either all are
\emph{applied} to Rocksdb or none of them are. Put another way we
should not apply a mini-batch from a ganged operation if any of the
other mini-batches have failed to commit to their RAFT log. This means
that the client can be stateless. The second observation is that
checking whether the other minibatches have committed is easy because
the different RAFT instances can communicate through shared memory on
the same machines unlike the more general case of distributed
participants in two phase commit.

We now describe our solution called ``ganged operations" in Cyclone. 
The first key problem we need to deal with is that the leaders for the different
RAFT instances to which the keys map can be located on different machines. To
avoid having a client co-ordinate different machines for a single operation, we
constrain the RAFT leader election algorithm for the different RAFT instances to
converge to the same machine. We do this by triggering re-elections for a RAFT
instance \emph{other} than that of the first physical log until it is on the same
machine as the \emph{first} physical log. This process is therefore resilient to
failed machines. RAFT requires that the leader have the most up to date log. To
ensure that the process converges we therefore ensure that the currently elected
leader brings a majority quorum up to date before triggering a reelection. We
also do not accept any requests from clients until all leaders are
co-located. We assume that no network partitions occur \emph{within a machine}
i.e. some RAFT instances are able to send and receive packets, while others are
not, a possibility we discount due to our single process design.

Next, we need to simultaneously inject the ganged operation into all
participating physical logs on the machine hosting the leaders. Clients always
dispatch batched writes to a fixed co-ordinating physical log/RAFT instance
(henceforth called the co-ordinator), which is then responsible for forwarding
the request to the participating logs. We make use of packet cloning primitives
to avoid making physical copies of the packet, generating indirect references
instead. The co-ordinator adds a ``nonce'' - a unique timestamp - to the
packet. In addition, the co-ordinator adds a view number to the packet
containing the term numbers of all the participating RAFT instances (read from
shared memory of the co-located leaders).  The event that applies a ganged
operation is described in the pseudocode of Figure~\ref{fig:ganged_ops}. We
defer the discussion of how we generate the nonce to later in this section. We
also assume that a unique barrier is allocated in shared memory for each ganged
operation. We discuss later how this is done without using dynamic allocation.

The complexity in Figure~\ref{fig:ganged_ops} arises from the need to implement
a shared memory barrier and handle failure during replication. Without failure,
Figure~\ref{fig:ganged_ops} is straightforward. A barrier is executed on all
RAFT instances. Once replication is complete on all participating physical logs,
a state indicated by all necessary bits being set in the barrier mask, the RAFT
instances simultaneously execute the ganged operation.

@@

The assumption in Figure~\ref{fig:ganged_ops} is that each ganged
operation is mapped to a unique barrier. We achieve this by using a
fixed piece of memory owned by the co-ordinator to hold the barrier
and write the nonce to it in order to indicate that the barrier is
active for the corresponding ganged operation. Participants watch for
the nonce to know when to execute the ganged operation barrier in
Figure~\ref{fig:ganged_ops}, while also monitoring the co-ordinators
's published view to detect the case where the ganged operation fails
to replicate on the co-ordinator's physical log.
